--- 
title: "Additional EDA"
author: "Misha Khan"
date: "2022-12-02"
output: html_document
---


```{r setup, message=FALSE, echo=FALSE, warning=FALSE}
library(data.table)
library(tidyverse)
library(dplyr)
library(plotly)
library(DT)
library(knitr)

# Initialize code chunk options
opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  eval=TRUE,
  echo = TRUE,
  cache = FALSE,
  fig.width = 7, 
  fig.align = 'center',
  fig.asp = 0.618,
  out.width = "700px")

source("crime_presentation_eda.R")
```

## Calculate proportion of crime per area
```{r, message=FALSE, echo=FALSE, warning=FALSE}
barplot(sort(v1, T)[1:10], las = 2, col = rainbow(12), cex.names= .7, main = "Top 10 Areas with Crime", xlab = "Area", ylab = "Count")
```
```{r, message=FALSE, echo=FALSE, warning=FALSE}
v1
```

This bar chart displays the proportion of crime in the top 10 areas in Los Angeles. Central and 77th Street are significantly higher than the rest.

<br>

## Examining the Crime in the Top 2 Areas in Los Angeles 
### Central Map
```{r, message=FALSE, echo=FALSE, warning=FALSE}
#Map color palette
central.pal <- colorFactor(c('red', 'yellow', 'blue', 'green', 'purple'), domain = central_top_crimes$`Crime`)

central_map <- 
  leaflet(central_top_crimes) %>%  
  addProviderTiles('CartoDB.Positron') %>% 
  addCircles(
    lat = ~Lat, 
    lng = ~Lon,
    label = ~paste0(central_top_crimes$`Crime`),
    color = ~ central.pal(central_top_crimes$`Crime`),
    opacity = 0.5,
    fillOpacity = 1,
    radius = 5
  ) %>%
  addLegend('bottomleft', 
            pal = central.pal,  
            values = central_top_crimes$`Crime`,
            title = 'Central Crime Map', 
            opacity = 1)
```
```{r, message=FALSE, echo=FALSE, warning=FALSE}
central_map
```

This is a map of top 5 crimes in Central area. There is a noticeable amount of blue dots which is the top reported crime, burglary from vehicle.

<br>

### 77th Street Map
```{r, message=FALSE, echo=FALSE, warning=FALSE}

st77.pal <- colorFactor(c('red', 'yellow', 'blue', 'green', 'purple'), domain = st77_top_crimes$`Crime`)

st77_map <- 
  leaflet(st77_top_crimes) %>%  
  addProviderTiles('CartoDB.Positron') %>% 
  addCircles(
    lat = ~Lat, 
    lng = ~Lon,
    label = ~paste0(st77_top_crimes$`Crime`),
    color = ~ st77.pal(st77_top_crimes$`Crime`),
    opacity = 0.5,
    fillOpacity = 1,
    radius = 5
  ) %>%
  addLegend('bottomleft', 
            pal = st77.pal,  
            values = st77_top_crimes$`Crime`,
            title = '77th Street Map', 
            opacity = 1)
```

```{r, message=FALSE, echo=FALSE, warning=FALSE}
st77_map
```

This is a map of top 5 crimes in 77th Street area. There is a noticeable amount of purple dots which is the top reported crime, vehicle stolen.

<br>




