---
title: "Los Angeles Crime"
author: "Misha Khan"
output:
  pdf_document:
    toc: yes
  html_document:
    toc: yes
    toc_float: yes
---

<br>

# **Introduction**
For my final project, I selected Los Angeles crime dataset starting from 2020. Because
Los Angeles is known for having higher crime rates than most populated cities, I thought
it would be interesting to examine local areas with the highest crime, most common crime, and victim demographics. 
The original dataset is a .csv file that comes from Los Angeles Open Data website. The dimensions are 586,295 observations and 28 variables that describe the date, time, crime, weapon, victim, and location. 

Data source: https://data.lacity.org/Public-Safety/Crime-Data-from-2020-to-Present/2nrs-mtv8 

<br>

# **Data Cleaning**
### Removing and Reformatting Columns
First, I renamed the columns and dropped 10 columns that were not relevant for analysis. Then, I reformatted date and time variables to extract month and year and to use them as variables. Since the year is not over, I removed 2022 for certain time related EDA/analysis.

```{r reformat, message=FALSE, echo=FALSE, warning=FALSE}
#source("pm566-midterm-final-copy.R")

library(stringr)
library(dplyr)
library(tidyverse)
library(rvest)
library(xml2)
library(dbplyr)
library(dplyr)
library(skimr) 
library(ggplot2)
library(data.table)
library(lubridate)
library(leaflet)
library(webshot)
library(rapportools)
library(lubridate)
library(dtplyr)

crimeraw <- read.csv("Crime_Data_from_2020_to_Present.csv")

#Rename the columns to be more descriptive
colnames(crimeraw) <- c('Record No',
                        'Date Reported',
                        'Date Occurred',
                        'Time Occurred',
                        'Area No',
                        'Area',
                        'Rpt Dist No', 
                        'Part', 
                        'Crime Code',
                        'Crime',
                        'MO', 
                        'Victim Age',
                        'Victim Sex',
                        'Victim Ethnicity',
                        'Premise Code',
                        'Premise', 
                        'Weapon Code',
                        'Weapon',
                        'Case Code',
                        'Case Status',
                        'Crime Code 1',
                        'Crime Code 2',
                        'Crime Code 3',
                        'Crime Code 4',
                        'Location',
                        'Cross Street',
                        'Lat',
                        'Lon'
                        )

#Drop unwanted columns
crimeraw <- select(crimeraw, -c('Rpt Dist No','Part', 'MO','Premise', 'Crime Code 1', 'Crime Code 2', 'Crime Code 3', 'Crime Code 4', 'Cross Street'))
```

```{r date, message=FALSE, echo=FALSE, warning=FALSE}
#Reformat date variables
crimeraw$`Date Reported` <- as.Date(crimeraw$`Date Reported`, format = "%m/%d/%Y")
crimeraw$`Date Occurred` <- as.Date(crimeraw$`Date Occurred`, format = "%m/%d/%Y")

#Change miltary to standard time
crimeraw$`Time Occurred` <- format(strptime(substr(as.POSIXct(sprintf("%04.0f", crimeraw$`Time Occurred`), 
                                 format="%H%M"), 12, 16), '%H:%M'), '%I:%M %p')
```

<br>

### Examining Coordinates
For coordinates, there were 2,266 coordinates that were (0,0) due to privacy reasons. I dropped these rows since the dataset is already large and it would not affect analysis.

```{r coord, message=FALSE, echo=FALSE, warning=FALSE}
#Drop (0,0) coordinates
sum(crimeraw$Lat == 0)
sum(crimeraw$Lon == 0)
crimeraw <- subset(crimeraw, Lat != 0, Lon != 0)
```

<br>

### Empty Cells and Negative Values
Victim Age had negative values, 0's (meaning not available), and an oddly high value of 120. For simplicity, I filtered age from 0-100 to exclude negative values. This step discarded 25 rows.
The variables Weapon, Victim Sex, and Victim Ethnicity had thousands of empty cells. For Victim Sex and Ethnicity, I replaced it with NA instead of dropping it. I assumed that the empty cells meant that the information was not given. As for weapon, I assigned the empty cells to "NONE" meaning that no weapon was used.

```{r age, message=FALSE, echo=FALSE, warning=FALSE}
#Examine Victim Age
summary(crimeraw$`Victim Age`)
crimeraw <-
  crimeraw %>%
  filter(`Victim Age` >= 0 & `Victim Age` <= 100)
```

```{r na1, message=FALSE, echo=FALSE, warning=FALSE}
#Replace empty cells with NA values
crimeraw$Weapon[is.empty(crimeraw$Weapon)] <- "NONE"
crimeraw$`Victim Sex`[is.empty(crimeraw$`Victim Sex`)] <- NA
crimeraw$`Victim Ethnicity`[is.empty(crimeraw$`Victim Ethnicity`)] <- NA
crimeraw$`Victim Age`[crimeraw$`Victim Age` == 0] <- NA
```

```{r na2, message=FALSE, echo=FALSE, warning=FALSE}
#Replace missing cells in Weapon Code with 999 to signify no weapon
crimeraw$`Weapon Code` <- crimeraw$`Weapon Code` %>% replace_na(999)
```

<br>

### After Cleaning
After data cleaning, the process removed 2,291 rows. The crimedat dataset now has 584,004 observations and 21 variables. For further analysis, I subsetted the data to top 5 areas with the highest crime because of how large the dataset is.

<br>

# **EDA**
## When does the most crime occur?
```{r, message=FALSE, echo=FALSE, warning=FALSE}
library(data.table)
library(tidyverse)
library(dplyr)
library(plotly)
library(DT)
library(knitr)

# Initialize code chunk options
opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  eval=TRUE,
  echo = TRUE,
  cache = FALSE,
  fig.width = 7, 
  fig.align = 'center',
  fig.asp = 0.618,
  out.width = "700px")

```

```{r, message=FALSE, echo=FALSE, warning=FALSE}
#Viz 1: Months with the Highest Crime in the Top 5 Areas
source("crime_presentation_eda.R")
month_crime = plot_ly(ncrimesdate, x = ~Month, y = ~n, color = ~Area, type = "bar")
month_crime <- month_crime %>% layout(yaxis = list(title = 'Count'), barmode = 'stack')
month_crime <- month_crime %>% layout(title = "Months with the Highest Crime in the Top 5 Areas")
month_crime_plotly <- ggplotly(month_crime)
```

#### Bar Chart {.tabset}
 
```{r, message=FALSE, echo=FALSE, warning=FALSE}
month_crime_plotly 
``` 
## {-}

Like mentioned before, 2022 is excluded from this analysis since the year is not over. The EDA above contains two outputs per area for the total number of reports in 2020 and 2021. Focusing on the top 5 areas with the highest crime, I saw that January to June there are a little under 10,000 reports. However, in the warmer months, the reports increase over 10,000 and then drop again in November and December.

<br>

## What are the most common crimes?
```{r, message=FALSE, echo=FALSE, warning=FALSE}
#Viz 2: Most Common Crimes
top_crime <- ncrimes %>% 
  plot_ly(x = ~Crime, y = ~n,
          type = 'scatter',
          mode = 'markers',
          color = ~Area,
          sizes = c(5, 70),
          marker = list(sizemode='diameter', opacity=0.5))
top_crime <- top_crime %>% layout(title = "Most Common Crime in Top Areas")
top_crime_plotly <- ggplotly(top_crime)
```

#### Scatter Plot {.tabset}

```{r, message=FALSE, echo=FALSE, warning=FALSE}
top_crime_plotly 
``` 

## {-}
There are two dots per area for each year (2020 and 2021). Across the plot, the highest points (blue and orange) are from the areas 77th Street and Central. There are over 1,500 reports of burglary from vehicle and vehicle stolen.

<br>

## What is the demographic breakdown of victims?
```{r, message=FALSE, echo=FALSE, warning=FALSE}
#Viz 3: Victim Age and Sex
vict_demo <-plot_ly(victim_demographics, x = ~victim_age_f, color = ~`Victim Sex`) %>%
  add_histogram()
vict_demo <- vict_demo %>% layout(title = "Histogram of Victim's Age and Gender",
                                  yaxis = list(
                                    title = "Count"
                                  ),
                                  xaxis = list(
                                    title = "Age"
                                  ))
vict_demo_plotly <- ggplotly(vict_demo)
```

```{r, message=FALSE, echo=FALSE, warning=FALSE}
#Viz 4: Victim Age and Ethnicity
vict_demo_eth <-plot_ly(victim_demographics, x = ~victim_age_f, color = ~`Victim Ethnicity`) %>%
  add_histogram()
vict_demo_eth <- vict_demo_eth %>% layout(title = "Histogram of Victim's Age and Ethnicity",
                                  yaxis = list(
                                    title = "Count"
                                  ),
                                  xaxis = list(
                                    title = "Age"
                                  ))
vict_demo_eth_plotly <- ggplotly(vict_demo_eth)
```

## {.tabset}

### Victim Sex

```{r echo=FALSE}
vict_demo_plotly
```  

### Victim Ethnicity

```{r echo=FALSE}
vict_demo_eth_plotly 
```

## {-}

As for victim demographics, younger women, older men, and Hispanics report more crime.

<br>

# **Conclusion**
From the EDA displayed, there are a higher number of crime reports in the warmer months than colder months. The highest number of reports is in August while the lowest number of reports is in Aprils. This would be interesting to look into as for what possible factors could be possibly contributing to that.

In the areas with the highest rate of crime, theft from vehicle and vehicle theft is the most common. The highest rates of crime occur in 77th Street and Central areas. 77th Street has 1,765 reports of vehicle stolen and Central has 1,660 reports of theft from vehicle in 2021.

From the histogram, there is a noticeable spike in women in their 20s reporting crimes compared to men. However after 30s, there are more men reporting crime than women. 
When examining victim's ethnicity, there are high number of reports by Hispanic residents in their 20s through 40s. Second highest number of victims is White then Black.

<br>
